Gene expression profiling of reproductive meristem identities in rice inflorescence

### 2016-02-01 ms.Rmd
New conclusion

This report describes changes in gene expression during the early stages of rice panicle development. Based on expression patterns of uncharacterized genes, transcription factors and genes with functions potentially related to hormone signalling, a switch between apical and axillary meristems was evident, followed by gradual changes during transitions between PBM and SM. This could be explained by the difference between the fates of the rachis meristem, which aborts after PBM differentiation [@Ikeda:2004tk], and the axillary meristems, which transition to determinate SM. The latter process may be comparable to the gradual meristem maturation during tomato inflorescence development [@Park:2012dg], rather than to the continual production of determinate floral meristems from the Arabidopsis inflorescence meristem. These results are expected to lead to mechanistic studies of the regulatory networks that establish gene expression in these tissues, as well as functional studies of the individual components, both of which will be vital for understanding the genetic factors influencing rice yield via panicle morphology.

>>>>

The molecular mechanisms controlling the meristematic activities that guide panicle development in rice remain largely uncharacterized. The dataset presented here addresses this by describing the transcriptome of specific meristem identities and the changes in gene expression during the early stages of panicle development. A switch between apical and axillary meristems was evident, followed by gradual changes during transitions between PBM and SM. This could be explained by the difference between the fates of the rachis meristem, which aborts after PBM differentiation [@Ikeda:2004tk], and the axillary meristems, which transition to determinate SM. The latter process may be comparable to the gradual meristem maturation during tomato inflorescence development [@Park:2012dg], rather than to the continual production of determinate floral meristems from the Arabidopsis inflorescence meristem. Promising uncharacterized genes were identified, including transcription factors with expression patterns similar to other regulators of panicle architecture, and others potentially related to hormone signalling such as the auxin-responsive genes in clusters `r getClusterName('dec', 'small')` and `r getClusterName('inc', 'small')`. Functional studies of these genes will reveal the mechanics of the regulatory networks that establish gene expression in meristems and their control by hormone signalling. Genes with expression specific to certain meristem types were also identified, which may additionally serve as valuable marker genes, for example in developmental studies of mutants affected in determination of reproductive meristem identity. Combined with current technology for the creation of mutants, this dataset promises to accelerate research on the molecular mechanisms controlling panicle development, which will be vital for the sustainable increase in grain yield required by population growth.
